---
title: "pull_census_data"
author: "Jenny Rempel"
date: "7/23/2021"
output: html_document
---

A few notes
* It will take some time to create statewide, block-group-level population & ACS estimates with spatial data. (See ~line 80.)

```{r packages & setup, include = FALSE}
# set knitr options
knitr::opts_chunk$set(echo = TRUE)

# load packages
library(tidyverse) # for tidy syntax
library(magrittr) # for more powerful tidy syntax
library(sp) # for spatial data work - vectors
library(sf) # for spatial data work - vectors
library(arcpullr) # for pulling data from ArcGIS servers
library(httr) # for building API calls
library(tigris) # for state and county boundaries
library(tidycensus) # for census data access
library(here) # for relative file paths
library(tictoc) # for timing how long code chunks take
library(mapview)# for quick mapping
library(terra) # for raster work
library(raster) # for raster work
library(areal) # for areal weighted interpolation
library(tmap) # for mapping
library(ggpubr) # for figures
library(rstatix)
library(ggthemes) # for figures
library(praise) # for positivity boosts :)

# define functions and global options
`%notin%` <- Negate(`%in%`)
global_crs <- st_crs(4269) # set global CRS
options(tigris_use_cache = TRUE)
options(stringsAsFactors = FALSE)
census_api_key("69b2a9fd058d6fdd18c67f3688ba9e909fb21f50") # my census API key
# to get your own, go to https://api.census.gov/data/key_signup.html

# build function that pulls CA boundary into global environment as sf object
get_ca_boundary <- function(crs) {
    states() %>%
      filter(NAME == "California") %>%
      rename(name = NAME) %>%
      dplyr::select(name) %>%
      st_transform(crs = crs)
}
```

```{r import california and minnesota, message = FALSE}
# import and clean California state boundary as a spatial object
ca_st_bound <- tigris::states() %>% # import all states
  filter(NAME == "California") %>% # filter for California
  st_transform(global_crs) %>% # transform to global CRS
  rename_with(tolower) # rename columns to lowercase

# import and clean Minnesota state boundary as a spatial object
mn_st_bound <- tigris::states() %>% # import all states
  filter(NAME == "Minnesota") %>% # filter for Minnesota
  st_transform(global_crs) %>% # transform to global CRS
  rename_with(tolower) # rename columns to lowercase

# confirm spatial objects are selected and projected correctly
ggplot() +
  geom_sf(data = ca_st_bound) +
  #geom_sf(data = mn_st_bound) +
  theme_minimal()
```

```{r load tigerline shapefiles}
## add block group info for mapping
block_group <- st_read("raw_data/tl_2010_06_bg10/tl_2010_06_bg10.shp")
block_group_orig <- block_group

# transform
block_group_t <- block_group %>%
  st_transform(global_crs)

# double check
st_crs(block_group_t)
```


```{r load dasymmetric map layer}
da_map = system.file("raw_data/capopgrid2020_100m_unrounded.tif", package = "spDataLarge")
das_map <- raster("raw_data/capopgrid2020_100m_unrounded.tif")
glimpse(das_map)
das_map
das_map <- setMinMax(das_map)
das_map@crs
image(das_map)
bb <- extent(-381105,540895, -400000, 520000)
r <- setExtent(das_map, bb, keepres=TRUE)
plot(r)
ggplot() + 
  geom_raster(data = r, aes(x=x,y=y))

plot(das_map)

tmap_leaflet(
  tm_shape(das_map) +
    tm_raster() +
    tmap_options(max.raster = c(plot = 97455400, view = 97455400)))

```


```{r pull census data, message = FALSE, cache = TRUE}
# pull census data variable tables
census_vars <- load_variables(year = 2020, dataset = "sf1") # 2010 census is most current right now; 2020 coming soon
census_vars_2000_sf3 <- load_variables(year = 2000, dataset = "sf3") # 
census_vars_2000_sf4 <- load_variables(year = 2020, dataset = "sf4")
census_vars_2000_sf2 <- load_variables(year = 2020, dataset = "sf2")
View(census_vars_2000_sf3)

acs_5y_vars <- load_variables(year = 2019, dataset = "acs5") # 2014-2019 5-year ACS estimates, down to the block group level
acs_1y_vars <- load_variables(year = 2019, dataset = "acs1") # 2019 1-yr ACS estimates

# simplify 2000 census for viewing
census_vars_2000_sf3_cleaner <- census_vars_2000_sf3 %>% 
  count(concept) 


# simplify ACS variables for viewing
acs_cleaner <- acs_5y_vars %>% 
  count(concept) 

census_cleaner <- census_vars %>% 
  count(concept) 

#old delete
ca_census_pop_tract <- ca_census_pop_blkgrp
saveRDS(ca_census_pop_tract, "intermediate_data/ca_census_pop_tract.rds")
rm(ca_census_pop_blkgrp)


P001001 # TOTAL POPULATION [1]
P003001 # 100-PERCENT COUNT OF THE POPULATION [1]
HCT012001 # MHI in 1999 (Median household income in 1999!!Total	)
P007003 # Total!!Hispanic or Latino!!White alone
P037015 # Total female bachelor's degree
P037015 # Total male bachelor's degree

#total_pop = "P001001" # TOTAL POPULATION [1]
#hundred_pct_count_popn = "P003001"
#mhi = "HCT012001", # MHI in 1999 (Median household income in 1999!!Total	)
#nh_white = "P007003", # Total!!Hispanic or Latino!!White alone
fem_bachelors = "P037015", # Total female bachelor's degree
male_bachelors = "P037015" # Total male bachelor's degree


# 2nd total popn estimate
tic()
ca_census_pop_blkgrp2 <- get_decennial(state = "CA",
  # county = "SET", # set county
  geography = "block group", # pull block group-level data
  # Note: we could go to block level
  variables = c(hundred_pct_count_popn = "P003001"# 100-PERCENT COUNT OF THE POPULATION [1]
  ), # variable code for total population
  year = 2000,                                         
  geometry = FALSE # include vector geometry %>%
) 
toc()


# NH white
tic()
ca_census_pop_blkgrp_nhwhite <- get_decennial(state = "CA",
  # county = "SET", # set county
  geography = "block group", # pull block group-level data
  # Note: we could go to block level
  variables = c(nh_white = "P007003"),
  year = 2000,                                         
  geometry = FALSE # include vector geometry %>%
) 
toc()
 
# MHI
tic()
ca_census_pop_blkgrp_mhi <- get_decennial(state = "CA",
  # county = "SET", # set county
  geography = "block group", # pull block group-level data
  # Note: we could go to block level
  variables = c(mhi = "HCT012001"),
  year = 2000,                                         
  geometry = FALSE # include vector geometry %>%
) 
toc()

# fem ed
tic()
ca_census_pop_blkgrp_fem_ed <- get_decennial(state = "CA",
  # county = "SET", # set county
  geography = "block group", # pull block group-level data
  # Note: we could go to block level
  variables = c(fem_bachelors = "P037015"),
  year = 2000,                                         
  geometry = FALSE # include vector geometry %>%
) 
toc()

# male ed
tic()
ca_census_pop_blkgrp_male_ed <- get_decennial(state = "CA",
  # county = "SET", # set county
  geography = "block group", # pull block group-level data
  # Note: we could go to block level
  variables = c(male_bachelors = "P037015"),
  year = 2000,                                         
  geometry = FALSE # include vector geometry %>%
) 
toc()

# save those files
saveRDS(ca_census_pop_blkgrp, "intermediate_outputs/ca_census_pop_blkgrp.rds")
saveRDS(ca_census_pop_blkgrp2, "intermediate_outputs/ca_census_pop_blkgrp2.rds")
saveRDS(ca_census_pop_blkgrp_nhwhite, "intermediate_outputs/ca_census_pop_blkgrp_nhwhite.rds")
saveRDS(ca_census_pop_blkgrp_mhi, "intermediate_outputs/ca_census_pop_blkgrp_mhi.rds")
saveRDS(ca_census_pop_blkgrp_fem_ed, "intermediate_outputs/ca_census_pop_blkgrp_fem_ed.rds")
saveRDS(ca_census_pop_blkgrp_male_ed, "intermediate_outputs/ca_census_pop_blkgrp_male_ed.rds")

## reformat tbl's
# total pop - JMLR redo this
ca_census_pop_blkgrp <- ca_census_pop_blkgrp %>%
  rename(total_pop = value) %>% # JMLR reformat
  dplyr::select(-c(NAME, variable))
glimpse(ca_census_pop_blkgrp)

# hundred pct count of pop
ca_census_pop_blkgrp2 <- ca_census_pop_blkgrp2 %>%
  rename(hundred_pct_count_popn = value) %>%
  dplyr::select(-c(NAME, variable))
glimpse(ca_census_pop_blkgrp2)

# NH white
ca_census_pop_blkgrp_nhwhite <- ca_census_pop_blkgrp_nhwhite %>%
  rename(nh_white = value) %>%
  dplyr::select(-c(NAME, variable))
glimpse(ca_census_pop_blkgrp_nhwhite)

# MHI
ca_census_pop_blkgrp_mhi <- ca_census_pop_blkgrp_mhi %>%
  rename(mhi = value) %>%
  dplyr::select(-c(NAME, variable))
glimpse(ca_census_pop_blkgrp_mhi)

# female education
ca_census_pop_blkgrp_fem_ed <- ca_census_pop_blkgrp_fem_ed %>%
  rename(fem_ed = value) %>%
  dplyr::select(-c(NAME, variable))
glimpse(ca_census_pop_blkgrp_fem_ed)

# male education
ca_census_pop_blkgrp_male_ed <- ca_census_pop_blkgrp_male_ed %>%
  rename(male_ed = value) %>%
  dplyr::select(-c(NAME, variable))
glimpse(ca_census_pop_blkgrp_male_ed)


#  Things I have observed - remale ed & male ed have the same numbers; that seems like a mistake.
# MHI has no data.
# % NH white
```


```{r older census data pull, message = FALSE, cache = TRUE}
# get CA block-group-level population estimates from the 2010 census
tic()
ca_census_pop_blkgrp <- get_decennial(state = "CA",
                                           # county = "SET", # set county
                                           geography = "block group", # pull block group-level data
                                              # Note: we could go to block level
                                           variables = c(total_popn = "P001001"), # variable code for total population
                                           geometry = FALSE # include vector geometry %>%
                                          ) 
toc()
  # Note: Attempting this statewide crashed R for me, so we may want to run this on Savio, iterate through by county, or use a pre-made block group file (blasphemy, I know)

# save that data
tic()
saveRDS(ca_census_pop_blkgrp, "intermediate_data/ca_census_pop_blkgrp.rds") # or could save a gpkg, gdb, or shapefiles (look into Arc options)
toc()


# get additional CA block-group-level population estimates from the 2010 census
tic()
ca_census_nh_white_pop_blkgrp <- get_decennial(state = "CA",
                                           # county = "SET", # set county
                                           geography = "block group", # pull block group-level data
                                              # Note: we could go to block level
                                           variables = c(white_alone = "P005003"), # # Total!!Not Hispanic or Latino!!White alone	

                                           geometry = FALSE # include vector geometry %>%
                                          ) 
toc()

# save that data
tic()
saveRDS(ca_census_nh_white_pop_blkgrp, "intermediate_data/ca_census_nh_white_pop_blkgrp.rds") # or could save a gpkg, gdb, or shapefiles (look into Arc options)
toc()
```


```{r join tabular census data to spatial files}

# reformat tabular data on pop
ca_census_pop_blkgrp <- ca_census_pop_blkgrp %>%
  rename(GEOID10 = GEOID,
    total_pop = value) %>%
  dplyr::select(-c(variable, NAME))

# reformat tabular data on race
ca_census_nh_white_pop_blkgrp <- ca_census_nh_white_pop_blkgrp %>%
  rename(GEOID10 = GEOID,
    nh_white_pop = value) %>%
  dplyr::select(-c(variable, NAME))

# join data
block_group_t_1 <- block_group_t %>%
  merge(ca_census_pop_blkgrp, by = "GEOID10")

block_group_t_2 <- block_group_t_1 %>%
  merge(ca_census_nh_white_pop_blkgrp, by = "GEOID10")

# check
glimpse(block_group_t_2)

# backup
saveRDS(block_group_t_2, "intermediate_data/block_group_t_2.rds")

# calculate % BIPOC
block_group_t_2 <- block_group_t_2 %>%
  mutate(pct_bipoc = ((total_pop - nh_white_pop)/total_pop))

# flag if block group is majority BIPOC
block_group_t_2 <- block_group_t_2 %>%
  mutate(maj_bipoc = ifelse(pct_bipoc >= 0.5, "YES", "NO"))

```

```{r which block groups do restoration sites fall within?}

# load restoration data
cdfw <- readRDS("intermediate_data/part1_point.rds")

# convert cdfw data frame to sf object
cdfw_sf <- cdfw %>%
  st_as_sf(
    coords = c("WorksiteLo", "WorksiteLa"),
    agr = "constant",
    crs = global_crs, # JMLR: why isn't this working???
    stringsAsFactors = FALSE,
    remove = TRUE
  )

# transform CRS for consistency
cdfw_sf_t <- cdfw_sf %>%
  st_transform(global_crs)

# find points within polygons
site_in_blkgrp <- st_join(cdfw_sf_t, block_group_t_2, join = st_within)

# check
glimpse(site_in_blkgrp)

# save that
saveRDS(site_in_blkgrp, "site_in_blkgrp.rds")
```

```{r quick map}
# quickly visualize all the restoration sites
ca_boundary <- get_ca_boundary(crs = global_crs)

# double check
st_crs(ca_boundary)

# plot
ggplot() +
  geom_sf(data = ca_boundary, fill = "lightgrey") +
  geom_sf(data = site_in_blkgrp, aes(color = maj_bipoc)) +
  theme_minimal()

# plot
ggplot() +
  geom_sf(data = ca_boundary, fill = "lightgrey") +
  geom_sf(data = site_in_blkgrp, aes(color = total_pop)) +
  theme_minimal()
```










```{r MN pull census data, message = FALSE, cache = TRUE}
# get MN block-group-level population estimates from the 2010 census
tic()
mn_census_pop_blkgrp <- get_decennial(state = "MN",
                                           # county = "SET", # set county
                                           geography = "tract", # pull block group-level data
                                           variables = "P001001", # variable code for total population
                                           geometry = TRUE) %>% # include vector geometry %>%
  select(-variable) %>% # drop variable column
  rename(tot_pop_census = value) %>% # rename value column to indicate what it captures
  rename_with(tolower) # rename columns to lowercase
toc()

# save that data
tic()
saveRDS(mn_census_pop_blkgrp, "intermediate_data/mn_census_pop_tract.rds") # or could save a gpkg
toc()






# Pull ACS data on MHI, % of population with a college education, % of population that is non-Hispanic white
ca_acs_blkgrp <- get_acs(state = "CA",
                                  #county = "SET", # set county
                                  geography = "block group", # pull block group-level data
                                    # for acs, we can only go to block group or place (ie. not block)
                                  variables = c("B01003_001", # total population
                                    "B02008_001" # white population (alone or in combination with other race)
                                    # add more
                                    ), 
                                  # variables codes for total population and
                                  
                                  geometry = FALSE) %>% # do not include vector geometry
  rename_with(tolower) %>% # rename columns to lowercase %>%
  select(-moe, -name) %>% # drop measure of error and name columns
  pivot_wider(names_from = "variable", values_from = "estimate") %>% # pivot from long to wide format
  rename(tot_pop_acs = B01003_001, # rename columns after pivot
         white_pop_acs = B02008_001) %>%
  mutate(white_prop_acs = white_pop_acs / tot_pop_acs,
         nonwhite_prop_acs = 1 - white_prop_acs) # add column giving proportion of population that is white
```

```{r buffer point for intersecting layers}
# Fresno, CA, is at 36.7338° N, 119.7846° W

```

```{r to do}
# 1. Pull sample census file (Eg. population)
# 2. 
```




# Newer analysis, 11/29/21


```{r load data}

# load geopackage of sites
study_points <- st_read("intermediate_data/study_points.gpkg") 
st_crs(study_points) # double check

# confirm which column is the unique identifier
n_distinct(study_points$obs_id) # unique identifier
n_distinct(study_points$comid) # not unique
nrow(study_points)

# load census spatial data
blk_grp_sf <- st_read("raw_data/census/shapefiles/CA_blck_grp_2000.shp") 

# transform to universal crs
blk_grp_transf <- blk_grp_sf %>%
  st_transform(global_crs)
st_crs(blk_grp_transf) # double check

# load census tabular data
census_part1 <- read_csv(here("raw_data", "census", "nhgis0001_ds147_2000_blck_grp.csv"),
                     col_names = TRUE)
census_part2 <- read_csv(here("raw_data", "census", "nhgis0002_ds147_2000_blck_grp.csv"),
                     col_names = TRUE)
census_part3 <- read_csv(here("raw_data", "census", "nhgis0002_ds152_2000_blck_grp.csv"),
                     col_names = TRUE)

```

```{r join tabular and spatial census data}
# columns of interest from part 2
# FXS001:      Total
# FYF001:      Not Hispanic or Latino >> White alone
# HD1013:      Male >> Bachelor's degree
# HD1029:      Female >> Bachelor's degree
# HF6001:      Median income in 1999

# i think i may need to pull more for bachelors, so i'm going to focus on NH white & MHI for now

# simplify
census_part1 <- census_part1 %>%
  dplyr::select(GISJOIN, FXS001)

census_part2 <- census_part2 %>%
  dplyr::select(GISJOIN, FYF001)

census_part3 <- census_part3 %>%
  dplyr::select(GISJOIN, HD1013, HD1029, HF6001)

census_tabular <- census_part1 %>%
  left_join(census_part2, by = "GISJOIN")

census_tabular <- census_tabular %>%
  left_join(census_part3, by = "GISJOIN")

# rename
census_tabular <- census_tabular %>%
  rename(total_pop = FXS001,
    nh_white = FYF001,
    male_bach = HD1013,
    fem_bach = HD1029,
    mhi = HF6001)

# join tabular to spatial data
blk_grp_ca <- blk_grp_transf %>%
  merge(census_tabular, by = "GISJOIN")

# add % bipoc column
blk_grp_ca <- blk_grp_ca %>%
  mutate(pctg_bipoc = ((total_pop - nh_white) / total_pop))

mapview(blk_grp_ca[1:1000,])
```



```{r buffer sites}
## test workflow on one point
# sample 1 point
point1 = study_points[2,]
point1

# before buffering, transform CRS from planar to projected
df_sf <- st_as_sf(point1) %>% st_transform(3488) #transform to NAD83(NSRS2007) / California Albers

# create a 1km buffer
df_sf_buff <- st_buffer(df_sf, 1000)

# test that
mapview(df_sf_buff)

# save
point1_buff <- df_sf_buff %>%
    st_transform(st_crs(study_points))


## replicate for all sites
# before buffering, transform CRS from planar to projected
df_sf <- st_as_sf(study_points) %>% 
  st_transform(3488) #transform to NAD83(NSRS2007) / California Albers

# create a 1km buffer
df_sf_buff <- st_buffer(df_sf, 1000)

# test that
mapview(df_sf_buff[1:100,])

# transform back to previous crs
study_points_buff <- df_sf_buff %>%
  st_transform(st_crs(study_points))

```

```{r quick point in polygon analysis}

# double check crs
st_crs(study_points)
st_crs(blk_grp_ca)

# find points within polygons
site_w_blkgrp_st_contain <- st_join(study_points, blk_grp_ca, join = st_within)
glimpse(site_w_blkgrp)
glimpse(blk_grp_ca)

site_w_blkgrp <-  site_w_blkgrp_st_contain
glimpse(site_w_blkgrp_st_contain)

# which sites dont join to a block group?
sites_not_in_blk_grp <- site_w_blkgrp %>% 
  filter(is.na(GISJOIN)) %>%   # 54 sites
  mapview()
 
# save that map
mapshot(sites_not_in_blk_grp, "intermediate_outputs/sites_not_in_blk_grp.html")

# save that file
saveRDS(site_w_blkgrp, "intermediate_data/site_w_blkgrp.rds")

# quick check if there's a difference
site_w_blkgrp_analysis <- site_w_blkgrp %>%
  filter(!is.na(GISJOIN)) %>%
  group_by(obs_type) %>%
  summarize(n = n(),
    mean_total_pop = mean(total_pop),
    mean_mhi = mean(mhi),
    mean_pctg_bipoc = mean(pctg_bipoc),
    na.rm = TRUE)
site_w_blkgrp_analysis


# plot simple histogram
ggplot(site_w_blkgrp) +
  geom_histogram(aes(x = total_pop, fill = obs_type), alpha = .6) + 
  labs(x = "Total Population") 
ggsave("intermediate_outputs/hist_total_pop.png", width = 8, height = 5)

# try a density plot with a jittered rug
ggplot(site_w_blkgrp) +
  geom_density(aes(x = total_pop, fill = obs_type), alpha = .6) +
  geom_rug(aes(x = total_pop, y = 0), 
    position = position_jitter(height = 0)) + 
  labs(x = "Total Population") 
ggsave("intermediate_outputs/density_total_pop.png", width = 8, height = 5)


# plot simple histogram
ggplot(site_w_blkgrp) +
  geom_histogram(aes(x = pctg_bipoc, fill = obs_type), alpha = .6) + 
  labs(x = "% BIPOC") 
ggsave("intermediate_outputs/hist_pctg_bipoc.png", width = 8, height = 5)

# try a density plot with a jittered rug
ggplot(site_w_blkgrp) +
  geom_density(aes(x = pctg_bipoc, fill = obs_type), alpha = .6) +
  geom_rug(aes(x = pctg_bipoc, y = 0), 
    position = position_jitter(height = 0)) + 
  labs(x = "% BIPOC") 
ggsave("intermediate_outputs/density_pctg_bipoc.png", width = 8, height = 5)

## JMLR question: Since this is a point analysis, should this all have been done in 
```

```{r areal interpolation set up for buffered sites}

#### first, test with single site ####
# identify areas of intersection
site_blk_grp_overlap_area_test<- st_intersection(x = point1_buff, 
                                         y = blk_grp_ca) %>% 
   mutate(site_blk_grp_intersect_area = st_area(.)) 

# view that
mapview(site_blk_grp_overlap_area_test)

# calculate overall area for each point
study_points_buff2 <- study_points_buff %>%
  mutate(buff_circle_area = st_area(study_points_buff))
summary(study_points_buff2$buff_circle_area) # area is always 3140157 m^2

# calculate % area for each
site_blk_grp_overlap_area <- site_blk_grp_overlap_area %>%
  mutate(pctg_intersect_area = site_blk_grp_intersect_area/3140157)

# simplify df
site_blk_grp_overlap_area2 <- site_blk_grp_overlap_area %>%
  dplyr::select(obs_id, GISJOIN, 
    total_pop, nh_white, mhi, pctg_bipoc, 
    site_blk_grp_intersect_area, pctg_intersect_area)

# apply those weights
site_blk_grp_overlap_area2 <-  site_blk_grp_overlap_area2 %>%
  mutate(total_pop_wtd = total_pop * pctg_intersect_area, # wtd = weighted
    nh_white_wtd = nh_white * pctg_intersect_area,
    mhi_wtd = mhi * pctg_intersect_area,
    pctg_bipoc_wtd = pctg_bipoc * pctg_intersect_area)

# sum across block groups associated with each point
site_blk_grp_overlap_area3 <- site_blk_grp_overlap_area2 %>%
    st_drop_geometry() %>%  # drop geometry for pivot
  group_by(obs_id) %>%
  summarize(total_pop_buff = sum(total_pop_wtd),
    nh_white_buff = sum(nh_white_wtd),
    mhi_wtd_buff = sum(mhi_wtd),
    pctg_bipoc_buff = sum(pctg_bipoc_wtd))






#### next, repeat for all sites ####

## prep sf objects for intersection
# transform in an attempt to address st_intersection assumption
study_points_buff_t <- study_points_buff %>%
  st_transform(3488) #transform to NAD83(NSRS2007) / California Albers

blk_grp_ca_t <- blk_grp_ca %>%
  st_transform(3488) #transform to NAD83(NSRS2007) / California Albers

# 10 of the block groups have non-valid geometries; i am cleaning them up simply with a 0-meter buffer - JMLR, ask Lucy
blk_grp_ca_t <- st_buffer(blk_grp_ca_t, dist = 0)

# check that it worked
test <- blk_grp_ca_t %>%
  mutate(valid_geom = st_is_valid(test))
test %>% filter(valid_geom == FALSE) # it worked

# identify areas of intersection
site_blk_grp_overlap_area <- st_intersection(x = study_points_buff_t, 
                                         y = blk_grp_ca_t) %>% 
   mutate(site_blk_grp_intersect_area = st_area(.)) 
  ### JMLR: warning message: attribute variables are assumed to be spatially constant throughout all geometries - dbl chk

# backup
site_blk_grp_overlap_area_orig <- site_blk_grp_overlap_area

# check that
mapview(site_blk_grp_overlap_area_orig[1:500,])

# calculate % area for each
site_blk_grp_overlap_area <- site_blk_grp_overlap_area %>%
  st_drop_geometry() %>%  # drop geometry for ease of processing
  mutate(pctg_intersect_area = site_blk_grp_intersect_area/3140157)


# add a check that the geometries sum to 100%
site_blk_grp_overlap_area_test <- site_blk_grp_overlap_area %>%
  group_by(obs_id) %>%
  summarize(summed_pctg_intersect_area = sum(pctg_intersect_area),
    summed_area = (sum(site_blk_grp_intersect_area)/3140157)) %>%
  ungroup()

# some of the geometries don't add up to 100% - look at those 713 projects
site_not_100pct_blk_grp_covered <- site_blk_grp_overlap_area_test %>%
  filter(as.numeric(summed_area) < 1)

# are they controls or actual projects?
site_not_100pct_blk_grp_covered2 <- study_points_buff %>% 
  left_join(site_not_100pct_blk_grp_covered, by = "obs_id")

site_not_100pct_blk_grp_covered2 %>%   filter(as.numeric(summed_area) < 1) %>% count(obs_type) 
  # 9 695 controls, 18 restoration sites

# map them
site_not_blk_grp_covered <- site_not_100pct_blk_grp_covered2 %>%
  filter(as.numeric(summed_area) < 1)
mapview(site_not_blk_grp_covered)
mapshot(site_not_blk_grp_covered, "intermediate_outputs/site_not_blk_grp_covered.html") # not saving; not sure why

## SESYNC JMLR: Decide: what to do with these sites out of place

# simplify df
site_blk_grp_overlap_area2 <- site_blk_grp_overlap_area %>%
  dplyr::select(obs_id, GISJOIN, 
    total_pop, nh_white, mhi, pctg_bipoc, 
    site_blk_grp_intersect_area, pctg_intersect_area)

# apply those weights
site_blk_grp_overlap_area2 <-  site_blk_grp_overlap_area2 %>%
  mutate(total_pop_wtd = total_pop * pctg_intersect_area, # wtd = weighted
    nh_white_wtd = nh_white * pctg_intersect_area,
    mhi_wtd = mhi * pctg_intersect_area,
    pctg_bipoc_wtd = pctg_bipoc * pctg_intersect_area)

# sum across block groups associated with each point
site_blk_grp_overlap_area3 <- site_blk_grp_overlap_area2 %>%
  group_by(obs_id) %>%
  summarize(total_pop_buff = sum(total_pop_wtd),
    nh_white_buff = sum(nh_white_wtd),
    mhi_wtd_buff = sum(mhi_wtd),
    pctg_bipoc_buff = sum(pctg_bipoc_wtd))

# backup
saveRDS(site_blk_grp_overlap_area3, "intermediate_data/site_blk_grp_overlap_area3.rds")

# remove units
site_blk_grp_overlap_area4 <- site_blk_grp_overlap_area3 %>%
  mutate(total_pop_buff_area = as.numeric(total_pop_buff),
    nh_white_buff_area = as.numeric(nh_white_buff),
    mhi_wtd_buff_area = as.numeric(mhi_wtd_buff),
    pctg_bipoc_buff_area = as.numeric(pctg_bipoc_buff)) %>%
  dplyr::select(obs_id, total_pop_buff_area, nh_white_buff_area, 
    mhi_wtd_buff_area, pctg_bipoc_buff_area)

# backup
saveRDS(site_blk_grp_overlap_area4, "intermediate_data/site_blk_grp_overlap_area4.rds")

# join back to summary data
study_points_w_ses <- study_points %>%
  left_join(site_blk_grp_overlap_area4, by = "obs_id")

# note to JMLR: i should do more checking to make sure this is working as expected

# export as geopackage
st_write(study_points_w_ses,
  dsn = "intermediate_data/study_points_w_ses.gpkg", 
  layer = "intermediate_data/study_points_w_ses.gpkg",
  driver = "GPKG")
# note this is in the wrong CRS
```

```{r quick t-test work}
# make a density plot
total_pop_dens <- ggdensity(study_points_w_ses, x = "total_pop_buff_area",
          add = "mean", color = "obs_type", fill = "obs_type",
          #rug = TRUE, 
          xlab = "Total population",
          ylab = "Density of restoration sites")
          #panel.labs = list(buy_type = 
                        #      c("Family Wealth", "LLC")))
total_pop_dens

stat.test <- study_points_w_ses %>% 
  t_test(total_pop_buff_area ~ obs_type) %>%
  add_significance()
stat.test

glimpse(study_points_w_ses)
```




*To Do*
/1. Download 2000 census appropriate data (tidycensus is not cooperating)
/1b. If needed, merge tabular to spatial data
/2. Create a sample point (eg. Fresno)
/3. Create a polygon that represents a 1km buffer around the sample point 
/4. Do areal weighted interpolation (ie. mean value) of the relevant data layer for that buffer**
/4b option. Pull census block group info for the BG that the point falls within
/5. Set it up to automate step 3 (ie. iterate through all points in a list and create buffered files)
/6. Set it up to automate step 4 (ie. iterate through aerial weighted interpolation)**
/6b option. Iterate thru simple option
/7. Upload file as gpckg**
**8. Pull updated education data**
**9. Get ready for stats: Run some fancier t-tests, pull pearson's coeff code**
**adjust the denominator for buffered sites that dont overlap census BGs**


*Question for group*
SESYNC JMLR: Decide: what to do with buffered sites that have some % of area not overlapping census block groups (or not overlapping BGs at all) --> decision: change denominator to be actual area that overlaps with census block groups.

*Next steps for group work*
1. Select subset of data with 
2. ID covariates & decide which attributes to keep
3. Control site pairwise matching
4. Run stats
5. Test for spatial autocorrelation; figure out if/how we need to account for temporal autocorrelation

